The symmetry problem of the \textit{Fermi} and eROSITA bubbles: A proof-of-concept study

### 2022-02-21 paper.tex
update

\begin{abstract}
A decade ago, the \textit{Fermi Gamma-Ray Space Telescope} reveals two large gamma-ray bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ on either side of the Galactic disk.\
The recent discovery of eROSITA bubbles also shows a symmetrically enhanced X-ray\
emission from the shells around the eROSITA bubbles.\
To this end, previous simulation-based works assumed that the jets are vertical to the Galactic disk;\
however, there is a lack of preferred orientation of jets with respect to the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium disk is an essential\
component for the symmetry Galactic bubbles.\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
the oblique jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the Galactic plane albeit\
the jet is at an angle $45^{\circ}$ with respect to the disc normal.\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a tangling contact discontinuity,\
which we interpret as the edge of the \textit{Fermi} bubbles.\
The nature of the \textit{Fermi} bubbles is essentially a turbulent and high-temperature ($\sim2$ keV)\
plasma in pressure balance with the external medium.
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between simulated and observed multiwavelength features\
releases the caveat given by earlier simulation-based studies:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

Early attempts \citep{Sarkar2015,Yang2017,Zhang2020}\
to model the symmetric \textit{Fermi} and eROSITA bubbles (collectively we call the Galactic bubbles)\
assumed that the jet is vertical to the Galactic plane. However,\
observation \citep{Gallimore2006} has proposed that\
there is a lack of preferred orientation of jets with respect to\
the rotation axis of a galactic plane (disc normal, hereafter).\
In contrast, a number of galaxies in which the jets are oblique to the disc normal\
(e.g. NGC 3079, \citealt{Cecil2001}; NGC 1052, \citealt{Dopita2015}),\
including galaxies in which the jets lie in the plane of the disc (e.g. IC 5063, \citealt{Morganti2015}).

To this end, this work introduces a dense thin disc of interstellar medium\
around a jet source to deflects and decelerates the inclined jets,\
in an attempt to resolve the symmetry problem of the Galactic bubbles.\

The main purpose of this study is to use three-dimensional\
special relativistic hydrodynamics simulations\
involving CR jet injections from the central SMBH in the Galaxy to investigate\
whether the \textit{inclined} jet scenario is able to produce\
the \textit{symmetric} Galactic bubbles that are\
consistent with the observed features, including the shape, the\
surface brightness, and the spectra of \textit{Fermi} bubbles \citep{Ackermann2014}\
and microwave haze \citep{Dobler_2008,PlanckCollaborationIX2013}.

@@

 The most interesting finding is that there is a innermost bubbles\
 (dashed box in left top panel of \Cref{fig__jetI5+ismSeed3-45deg})\
 extending out from the Galactic center on either side of the thin disk.\
 The innermost bubbles are cold (1-10 eV), dense ($10^{-4}$--$10^{-2}$ cm$^{-3}$),\
 and underpressured with respect to the turbulent plasma.\
 Moreover, the close-up view of profiles around $z=3.7$ kpc (right column in \Cref{fig__profile})
 demonstrates that there is a sharp pressure jump and dense shell at $z=3.61$ kpc,\
 indicating that the innermost bubbles are an expanding reverse shock.\
 We also find that the turbulent plasma is bracketed between the downstream of reverse shock\
 and of the outermost forward shock, thus heating the turbulent plasma up considerably.

@@

  First, the broad agreement between simulated and observed X-ray maps hints that\
  the full vertical extent of the eROSITA bubbles can be properly formed by an inclined jet within a thin disk\
  of dense interstellar medium.\

@@

\section{Conclusions}
\label{Conclusions}
In this work, we introduce a thin, dense disk composed of clumpy interstellar medium\
to divert the inclined jet at an angle $45^{\circ}$\
to the disk normal in the past energetic event of the central SMBH.

@@

\item The Galactic bubbles are nearly symmetric about the Galactic plane albeit the jet is at an angle $45^{\circ}$\
      with respect to the disc normal.\
      The broad agreement between simulated and observed multiwavelength features\
      provides supporting evidence for the inclined jet scenario, and also\
      releases the caveat given by earlier simulation-based studies:\
      the jets shall be vertical to the disc normal.
\item The randomness of clumpy disk has a insignificant impact on the overall dynamic of the Galactic bubbles.\
      The development of the reverse shocks (the innermost bubbles) are always associated with the dense disc,\
      without the disc, the outermost bubbles and the turbulent plasma will be oblique,\
      indicating the dense disc is crucial for\
      the symmetry of the Galactic bubbles and for the innermost bubbles formation.
\item The edge of the eROSITA bubbles is a forward shock,\
      originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
      and significantly stretched by the stratified atmosphere afterwards.\
      The simulated eROSITA bubbles are not as limb-brightened as the observed one.\
      A possibility to enhance the X-ray emission is to include shock-accelerated CRs near the shock,\
      in which CRs could increase the compressibility of the fluid.\
\item Followed by the forward shock is a tangling contact discontinuity,\
      which we interpret as the edge of the \textit{Fermi} bubbles.\
      The nature of the \textit{Fermi} bubbles is essentially a turbulent and high-temperature ($\sim2$ keV)\
      plasma in pressure balance with the external medium. Associated with the Galactic magnetic field\
      will lead to the stochastic acceleration of CRe, and probably balanced with the IC and synchrotron cooling.\
      We will investigate the competition between stochastic acceleration and radiative cooling in a future work.
\item According to the microwave and leptonic gamma-ray spectra, the best-fitting\
      CRe power-law index is found to be 2.4.
\item The same leptotonic CRe can simultaneously\
      account for the \textit{Fermi} bubbles and haze emission suggested that\
      they are physically related around the GC,\
      and that the magnetic fields within the bubbles are close to the exponentially distributed\
      Galactic magnetic field.
\end{itemize}

>>>>

\begin{abstract}
A decade ago, the \textit{Fermi Gamma-Ray Space Telescope} reveals two large bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ on either side of the Galactic center (GC).\
The recent discovery of eROSITA bubbles also shows a symmetrically enhanced X-ray\
emission from the shells around the eROSITA bubbles.\
To this end, previous simulation-based works assumed that the jets are vertical to the Galactic disk;\
however, there is a lack of preferred orientation of jets with respect to the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium disk is an essential\
component for the symmetry \textit{Fermi} and eROSITA bubbles (collectively, Galactic bubbles).\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
the oblique jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the GC albeit\
the bipolar jets are at an angle $45^{\circ}$ with respect to the disc normal.\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a tangling contact discontinuity,\
which we interpret as the edge of the \textit{Fermi} bubbles.\
The nature of the \textit{Fermi} bubbles is essentially a turbulent and high-temperature ($\sim2$ keV)\
plasma in pressure balance with the external medium.
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between simulated and observed multiwavelength features\
releases the caveat given by earlier simulation-based studies:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

Early attempts \citep{Sarkar2015,Yang2017,Zhang2020}\
to model the symmetric \textit{Fermi} and eROSITA bubbles (collectively we call the Galactic bubbles)\
assumed that the jets are vertical to the Galactic plane. However,\
observation \citep{Gallimore2006} has proposed that\
there is a lack of preferred orientation of jets with respect to\
the rotation axis of a galactic plane (disc normal, hereafter).\
In contrast, a number of galaxies in which the jets are oblique to the disc normal\
(e.g. NGC 3079, \citealt{Cecil2001}; NGC 1052, \citealt{Dopita2015}),\
including galaxies in which the jets lie in the plane of the disc (e.g. IC 5063, \citealt{Morganti2015}).

To this end, this work introduces a dense thin disc of interstellar medium\
around a jet source to deflects and decelerates the oblique jets,\
in an attempt to resolve the symmetry problem of the Galactic bubbles.\

The main purpose of this study is to use three-dimensional\
special relativistic hydrodynamics simulations\
involving CR jet injections from the central SMBH in the Galaxy to investigate\
whether the \textit{oblique} jet scenario is able to produce\
the \textit{symmetric} Galactic bubbles that are\
consistent with the observed features, including the shape, the\
surface brightness, and the spectra of \textit{Fermi} bubbles \citep{Ackermann2014}\
and microwave haze \citep{Dobler_2008,PlanckCollaborationIX2013}.

@@

 The most interesting finding is that there is a innermost bubbles\
 (dashed box in left top panel of \Cref{fig__jetI5+ismSeed3-45deg})\
 extending out from the GC on either side of the thin disk.\
 The innermost bubbles are cold (1-10 eV), dense ($10^{-4}$--$10^{-2}$ cm$^{-3}$),\
 and underpressured with respect to the turbulent plasma.\
 Moreover, the close-up view of profiles around $z=3.7$ kpc (right column in \Cref{fig__profile})
 demonstrates that there is a sharp pressure jump and dense shell at $z=3.61$ kpc,\
 indicating that the innermost bubbles are an expanding reverse shock.\
 We also find that the turbulent plasma is bracketed between the downstream of reverse shock\
 and of the outermost forward shock, thus heating the turbulent plasma up considerably.

@@

  First, the broad agreement between simulated and observed X-ray maps hints that\
  the full vertical extent of the eROSITA bubbles can be properly formed by an oblique jet within a thin disk\
  of dense interstellar medium.\

@@

\section{Conclusions}
\label{Conclusions}
In this work, we introduce a thin, dense disk composed of clumpy interstellar medium\
to divert the oblique jets at an angle $45^{\circ}$\
to the disk normal in the past energetic event of the central SMBH.

@@

\item The Galactic bubbles are nearly symmetric about the Galactic plane albeit\
      the bipolar jets are at an angle $45^{\circ}$\
      with respect to the disc normal.\
      The broad agreement between simulated and observed multiwavelength features\
      provides supporting evidence for the oblique jet scenario, and also\
      releases the caveat given by earlier simulation-based studies:\
      the jets shall be vertical to the disc normal.
\item The randomness of clumpy disk has a insignificant impact on the overall dynamic of the Galactic bubbles.\
      The development of the reverse shocks (the innermost bubbles) are always associated with the dense disc,\
      without the disc, the outermost bubbles and the turbulent plasma will be oblique,\
      indicating the dense disc is crucial for\
      the symmetry of the Galactic bubbles and for the innermost bubbles formation.
\item The edge of the eROSITA bubbles is a forward shock,\
      originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
      and significantly stretched by the stratified atmosphere afterwards.\
      The simulated eROSITA bubbles are not as limb-brightened as the observed one.\
      A possibility to enhance the X-ray emission is to include shock-accelerated CRs near the shock,\
      in which CRs could increase the compressibility of the fluid.\
\item Followed by the forward shock is a tangling contact discontinuity,\
      which we interpret as the edge of the \textit{Fermi} bubbles.\
      The nature of the \textit{Fermi} bubbles is essentially a turbulent and high-temperature ($\sim2$ keV)\
      plasma in pressure balance with the external medium. Associated with the Galactic magnetic field\
      will lead to the stochastic acceleration of CRe, and probably balanced with the IC and synchrotron cooling.\
      We will investigate the competition between stochastic acceleration and radiative cooling in a future work.
\item According to the microwave and leptonic gamma-ray spectra, the best-fitting\
      CRe power-law index is found to be 2.4.
\item The same leptotonic CRe can simultaneously\
      account for the \textit{Fermi} bubbles and haze emission suggested that\
      they are physically related around the GC,\
      and that the magnetic fields within the bubbles are close to the exponentially distributed\
      Galactic magnetic field.
\end{itemize}
